---
title: "rainbow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rainbow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rainbow)
```

## The `rivers_data` dataset

The dataset `rivers_data` contains measurements from three synthetic
stream sites: **Arikaree River**, **Caribou Creek**, and **Lewis Run**.
For each sampling event, the data record:

- **Site** – which stream the sample comes from  
- **Nitrate** – nitrate concentration in µmol/L  
- **Temperature** – water temperature in °C  
- **Dissolved oxygen** – in mg/L  
- **Conductance** – electrical conductance in µS/cm  
- **Turbidity** – a measure of water clarity (FNU)  
- **Elevation** – site elevation in metres  

Although the data are synthetic, they are based on patterns seen in real
studies and are intended for exploring basic data-visualisation and app
development ideas.

**DESCRIBING THE FUNCTION**

## The `my_first_plot()` function

The main example function in the package is `my_first_plot()`.  
It:

- Uses the `rivers_data` dataset shipped with the package  
- Produces a `ggplot2` boxplot of nitrate concentration by site  
- Applies a clean minimal theme so the focus stays on the data

Because `my_first_plot()` returns a `ggplot` object, it can be further
customised by users in their own analysis, for example by changing the
labels, colours, or theme.

This package contains one small synthetic dataset, rivers_data, and one demo plotting function, my_first_plot(). The dataset records nitrate concentration and common water-quality covariates at three stream sites. The plot function shows a boxplot of nitrate by site. A simple Shiny app is included to interactively filter sites and view a summary.

## Exploring the `rivers_data` dataset

```{r}
#We can take a quick look at the structure of the dataset.
str(rivers_data)
```

## Exploring the `rivers_data` dataset

```{r}
print(rainbow::my_first_plot())
```

**HOW TO USE THE PACKAGE**


A typical workflow with **rainbow** might be:

1. Load the package with `library(rainbow)`.  
2. Read the help pages for `rivers_data` and `my_first_plot()` to
   understand the variables and the example plot.  
3. Use `my_first_plot()` to quickly visualise nitrate differences between
   the three sites.  
4. Open the Shiny app with `launch_app()` to explore the data
   interactively and show results to others.

